---
title: "Domestic Violence Analysis in NSW"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents


```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
```

```{r libraryecho = FALSE, results='hide',message=FALSE, warning=FALSE}
library(dplyr)
library(arm)
library(ggplot2)
library(scales)
library(reshape2)
library(sqldf)
library(ggmap)
library(caret)
library(glmnet)
library(lme4)
library(jtools)
library(ggstance)
library(kableExtra)
library(knitr)
library(dplyr)

```



```{r Load Data, warning=FALSE}
#Loading data into datafeame
DV_NSW <- read.csv("data/DV_NSW_by_LGA.csv")

DV_Labels <- read.csv("data/labels.csv")

NSW_LGA <- read.csv("data/NSW_LGA.csv")


```

# Introduction 
&nbsp;&nbsp;&nbsp;Domestic violence (DV) is dark stain in modern society. DV can be defined as any form of mistreatment including physical, phycological, financial, sexual take place within domestic circumstances perpetrated by member of intimate relationship. Any member of household can be victim of DV, while women and children are most vulnerable to this crime. The World Health Origination(WHO) has described the violence against women is a global problem of public health which need immediate attention[1]. Australia is not Immune to this crime, especially NSW suffer with this crime heavily. Hance it is very important the understand the nature of this crime and identify the factors having most influence. In this project we shall try find these factors and try to prove following hypothesis by implementing a linear model.

**Null Hypothesis ($H_{0}$):** There is no relationship between income, education and employment with DV.\
**Alternative Hypothesis ($H_{A}$):** There is strong correlation between income, education and employment with DV.


\vspace{2mm}
# Data and methodology 
&nbsp;&nbsp;&nbsp;The dataset contains 3 CSV file. The $1^{st}$ CSV file "DV_NSW_by_LGA.csv" contains the count of DV crime in NSW LGA from Jan-99 to Dec-15. The $2^{nd}$ CSV file "NSW_LGA.csv" contain 2011 census data of NSW LGA. In the $2^{nd}$ CSV file there are around 8K features on 2011 NSW LGA census data. The features are in encoded form, where the description of the features is found in the $3^{rd}$ CSV file "labels.csv". At first we have performed some exploratory analysis to understand the dataset. We have generated some visualization to show the trend of the crime. From the visualization it is clear the trend is upward. To identify the features contributes most towards the crime we have implemented Forward stepwise Feature selection. 

&nbsp;&nbsp;&nbsp;**Forward stepwise Feature selection** is a computationally efficient algorithm to select the best features that produces best outcome. The algorithm starts with a null model and add predictors one by one until all predictors are tried. At each step, the predictor that give the largest improvement on the model is added to the final predictor list. The steps of the algorithm are as follows.

$1.$ Let, $M_{0}$ be a null model, which has no predictor.

$2.$ For $k=0$, .. , $p-1:$

> $(a)$ Consider all $p-k$ models that supplement the predictors in $M_{k}$ with one added predictor.\
> $(b)$ Chose the *best* between these $p-k$ models, and name it $M_{k+1}$. Here *best* is considered as the model having minimum RSS or maximum $R^2$.

$3.$ Select the best model from $M_{0}$, .., $M_{p}$ using any one of these metric cross-validated prediction error or $C_{p}$ $(AIC)$ or $(BIC)$ or adjusted $R^2$.

Using the features identified we shall build a linear model to predict DV.

\newpage
# Results and implications

&nbsp;&nbsp;&nbsp;The plot below visualizes yearly frequencies of all across NSW. From the plot it is clear the crime rate has an upward trend.

```{r Yearly Frequency, fig.height = 4, fig.width = 8, fig.align = "center", warning=FALSE}
#
DV_NSW1 <- melt(DV_NSW, id.vars="LGA") %>%
  dplyr::rename(Month = variable,
                Count = value)

DV_NSW1$Month <- format(as.Date(paste("1-",gsub("\\.", "-", DV_NSW1$Month),sep=""), "%d-%b-%y"),"%b-%y")

DV_NSW1$Year <- format(as.Date(paste("1-",gsub("\\.", "-", DV_NSW1$Month),sep=""), "%d-%b-%y"),"%Y")

#Creating a dataframe with yearly total count.
DV_NSW2 <- sqldf('SELECT Year, SUM(Count) AS YearCount FROM DV_NSW1 GROUP BY Year')

p1 <- ggplot(data=DV_NSW2, aes(x=Year, y=YearCount)) +
  labs( y = "Frequency", title = "Frequency by Year") +
  geom_bar(stat="identity", aes(fill = YearCount)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10, vjust = -.25, angle = 90, hjust = 1)) +
  theme(legend.position = "none")
p1


```

&nbsp;&nbsp;&nbsp;The plot below visualizes the frequency of top 10 LGA in descending order. Blacktown LGA outcast all other LGA by very clear margin. The concern authority might need to pay extra attention on Blacktown.


```{r Top 10 LGA Frequency, fig.height = 4, fig.width = 8, fig.align = "center", warning=FALSE}
#Creating a dataframe with yearly count of top 10 LGA.
DV_NSW3 <- sqldf('SELECT LGA, SUM(Count) AS YearCount FROM DV_NSW1 GROUP BY LGA order by YearCount DESC limit 10')

p2 <- ggplot(data=DV_NSW3, aes(x = reorder(LGA, -YearCount), y = YearCount)) +
  labs(x = "LGA", y = "Year Count", title = "Top 10 LGA") +
  geom_bar(stat="identity", aes(fill = YearCount)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10, vjust = -.25, angle = 90, hjust = 1)) +
  theme(legend.position = "none")

p2


```

\newpage
&nbsp;&nbsp;&nbsp;The plot below visualizes the frequency timeseries of top 4 LGA. The timeseries plot shows normal distribution of the frequency, no sudden abrupt spike or drop of the frequency. 

```{r Top 4 LGA Timeseries, fig.height = 7, fig.width = 10, fig.align = "center", warning=FALSE}

#DV_NSW1_Top <- dplyr::filter(DV_NSW1, LGA %in% c('Campbelltown', 'Blacktown', 'Penrith', 'Wyong', 'Liverpool', 'Fairfield', 'Parramatta', 'Newcastle', 'Lake Macquarie', 'Wollongong'))
#Creation dataframe with top 4 LGA data
DV_NSW4 <- subset(DV_NSW1, LGA == "Blacktown" | LGA == "Campbelltown" | LGA == "Penrith" | LGA == "Liverpool")


DV_NSW4$Month <- as.Date(paste("1-",DV_NSW4$Month,sep = ""),"%d-%b-%y")
DV_NSW4$Year <- as.numeric(DV_NSW4$Year)


p3 <- ggplot(data = DV_NSW4, aes(x = Month, y = Count,  color=LGA)) +
  geom_point(alpha = 0.01) + geom_line(aes(group = 1)) +
  facet_wrap(~LGA, ncol=1, scales = "free_x") +
  labs( title = "Top 4 LGA Time Series") +
  scale_y_continuous() +
  scale_x_date(labels = date_format("%b-%y"), breaks = date_breaks("5 year")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 6)) +
  theme_bw()+
  theme_minimal() +
  theme(axis.text.x.bottom = element_text(angle = 90, vjust = 0.5, size = 6)) +
  theme(axis.text.y = element_text(size = 6)) +
  theme(legend.position = "none")

p3

```


```{r 2011, message=FALSE,warning=FALSE}
# As the  ABS census data packs contains only data of 2011, as shall use only 2011 LGA DV data for model building purpose.
DV_NSW_2011 <- filter(DV_NSW1, Year == '2011')

#Creating dataframe with 2011 total count for each LGA
DV_NSW_2011_sum <- DV_NSW_2011 %>%
  group_by(LGA) %>%
  dplyr::summarise(DV_Count_2011 = sum(Count)) %>%
  arrange(LGA)
#Cleaning the LGA name by removing (C) and (A) form LGA name
NSW_LGA$label <- gsub(" \\(C)", "", NSW_LGA$label)
NSW_LGA$label <- gsub(" \\(A)", "", NSW_LGA$label)

#Joing the DV data with ABS census data
DV_NSW_2011_all <- DV_NSW_2011_sum %>%
  left_join(NSW_LGA %>%
              dplyr::rename(LGA = label)
              )

#Encoding the LGA name to numaric value
DV_NSW_2011_all$LGA <- as.numeric(as.factor(as.character(DV_NSW_2011_all$LGA)))

#Removing the columns region_id from from the dataframe
DV_NSW_2011_all <- DV_NSW_2011_all[-3]

#Scaling the data to standard form
for(i in 4:ncol(DV_NSW_2011_all)){
  current.f <- colnames(DV_NSW_2011_all)[i]
  DV_NSW_2011_all$current.f <- scale(DV_NSW_2011_all[,current.f])
}


```

\
&nbsp;&nbsp;&nbsp;After applying the Forward stepwise Feature selection algorithm, the features selected by the algorithm are listed in the table below. From the feature list we can see, no features on income education and employment as our hypothesis are selected by the algorithm. Therefore, we are manually adding feature B115, B126, B5503 and B2847 in the predictor list covering these aspects. With this combined predictor list, we shall build a linear model.

```{r forward stepwise, message=FALSE,warning=FALSE}

selectFeature <- function(train, test, cls.train, cls.test, features) {
  ## identify a feature to be selected
  curr.min.mse <- Inf
  selected.i <- NULL
  for(i in 1:ncol(train)) {
    current.f <- colnames(train)[i]

    # Skip the features already selectd and the response variable DV_Count_2011
    if(current.f %in% c(features, "DV_Count_2011")) { next }
    model <- lm(data=train[,c(features, current.f, "DV_Count_2011")], DV_Count_2011 ~ .)
    #Checking the MSE with only the features already selected plus ith feature and DV_Count_2011
    test.mse <- mean((cls.test - predict.lm(model, test[,c(features, current.f, "DV_Count_2011")])) ^ 2)

    if(test.mse < curr.min.mse) {
        curr.min.mse <- test.mse
        selected.i <- colnames(train)[i]
    }
  }
  return(selected.i)
}


set.seed(1)
inTrain <- createDataPartition(DV_NSW_2011_all$DV_Count_2011, p = .6)[[1]]

# Split the training and test data
train <- DV_NSW_2011_all[ inTrain,]
test  <- DV_NSW_2011_all[-inTrain,]

# Select the first feature with min error
features <- c()
curr.min.mse <- Inf

# Select the first feature
for(i in 1:ncol(train)) {
  #DV_Count_2011 is response variable y, hance not considering as feature in the dataset
  if(colnames(train)[i] == "DV_Count_2011") { next }
  #Checking the MSE with only the ith feature and DV_Count_2011
  model <- lm(data=train[,c(colnames(train)[i], "DV_Count_2011")], DV_Count_2011 ~ .)
  # Calculate the MSE
  test.mse <- mean((test$DV_Count_2011 - predict.lm(model, test[, c(colnames(train)[i], "DV_Count_2011")])) ^ 2)
  #Finding the feature with minimum MSE in the dataset
  if(test.mse < curr.min.mse) {
    curr.min.mse <- test.mse
    features <- colnames(train)[i]
    }
  }

#The above loop returns the feature that has min MSE
# Now finding the next nine features
for (i in 1:9) {
  selected.i <- selectFeature(train, test, train$DV_Count_2011, test$DV_Count_2011, features)

  # add the best feature from current run
  features <- c(features, selected.i)
}

DV_Labels_Sel <- DV_Labels %>%
  dplyr::filter(Sequential %in% features) %>%
  dplyr::select(Sequential, Long)

#The features are given as a long string concatened with _. Replacing _ with space for beatter readabelity
DV_Labels_Sel$Long <- gsub("_", " ", DV_Labels_Sel$Long)

kable(DV_Labels_Sel) %>%
  kable_styling(full_width = F) %>%
  row_spec(0, bold = T) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "30em")


```


&nbsp;&nbsp;&nbsp;We are building a linear model with the selected features. The summary of the linear model as presented below.

```{r LM model, fig.height = 4, fig.width = 6, fig.align = "center", warning=FALSE}


lm_fit <- lm(DV_Count_2011 ~ B115 + B126 + B5503+ B2847+ B4849 + B4702 + B2252 + B4258 + B2894 + B1758 + B3851 + B7452 + B2409 + B1390,DV_NSW_2011_all)

summary(lm_fit)


```

&nbsp;&nbsp;&nbsp;The model summary shows, the features added manually B115, B126, B5503 and B2847 has P-Value greater than 0.05, hance these are not a good predictor in the model. Therefore, we cannot reject $H_{0}$, i.e. There is no relationship between income, education and employment with DV. The features selected by Forward stepwise Feature selection algorithm are the best predictor to predict DV. The regression summary we can see these features have very low P-Value indicating the features having very strong relationship with the response variable. This fact is marked by *** in the summary. Also, the Multiple R-squared of the model returned 0.9831, indicating the predictors can explain 98.12% of variance from the model.\


```{r LM Model Plot, fig.height = 4, fig.width = 6, fig.align = "center", warning=FALSE}
plot(lm_fit)


```
\newpage
&nbsp;&nbsp;&nbsp;One of the best ways to present finding from a model is to generate plot_summs from jtools library. Using the plot, we can show the best predictor’s coefficient estimations are normally distributed as presented in plot below.

```{r LM Model Plot SUMMS, fig.height = 4, fig.width = 6, fig.align = "center", warning=FALSE}


plot_summs(lm_fit, scale = TRUE, plot.distributions = TRUE, inner_ci_level = .9)

```

&nbsp;&nbsp;&nbsp;Another useful visualization is effect_plot to show the relationship between predictor and response variable. The predictors with low P-Value are visualized below. All these plot display very strong connection with the response variable.


```{r LM Model effect Plot, fig.height = 3, fig.width = 6, fig.align = "center", warning=FALSE}

effect_plot(lm_fit, pred = "B4849", interval = TRUE, plot.points = TRUE)
effect_plot(lm_fit, pred = "B4702", interval = TRUE, plot.points = TRUE)
effect_plot(lm_fit, pred = "B2252", interval = TRUE, plot.points = TRUE)
effect_plot(lm_fit, pred = "B4258", interval = TRUE, plot.points = TRUE)
effect_plot(lm_fit, pred = "B2894", interval = TRUE, plot.points = TRUE)
effect_plot(lm_fit, pred = "B1758", interval = TRUE, plot.points = TRUE)
effect_plot(lm_fit, pred = "B3851", interval = TRUE, plot.points = TRUE)
effect_plot(lm_fit, pred = "B1390", interval = TRUE, plot.points = TRUE)



```


# Conclusion

&nbsp;&nbsp;&nbsp;DV is a social problem in NSW. The data we have used in this project, clearly shows the level of crime does not have a downward trend. Policy maker and law enforcing agencies need to have more clear insight on the crime, so that they can come up with a policy to reduce this crime. This project can add a very good value in this regard. We have shown the trend of the crime and the area where the crime is happening the most. We have also identified some key predictors and build a regression model that can predict the crime very with very high level of confidence. The data we have worked on is not enough to build a complete model. We had only census data of 2011. In future we shall add more census and DV data and try to build a complete model that can predict DV with higher accuracy.